\bf DOGE's Downsizing, Can AI Read the Reddit Room?

### 2025-04-10 code/ml_llm/R/assign_3.qmd
3rd draft of assing 3

---
title: "DOGE's Downsizing, Can AI Read the Reddit Room?"
format:
  jasa-pdf:
    keep-tex: true  
    journal:
      blinded: false
  jasa-html: default
date: last-modified
author:
  - name: Kevin Linares
    affiliations:
      - name: University of Maryland
  - name: Felix Baez-Santiago
  - name: Aria Lu
  - name: Gloria Zhou
    affiliations:
      - name: University of Michigan
abstract: |
  We investigated public sentiment on Reddit regarding the Department of Government Efficiency's federal workforce reduction by classifying 400 labeled comments (28% favor, 18% neutral, 54% oppose) using supervised and unsupervised Large Language Models (LLMs). It turned out that supervised models showed low to moderate success, particularly with "oppose" comments, but struggled with "favor" and "neutral" stances. Similarly, LLMs best intent identified "oppose" sentiment but exhibited low precision for "favor" and "neutral." These findings highlight the challenges of accurately gauging nuanced public opinion on government policy changes through automated methods on social media data.

@@

The newly formed Department of Government Efficiency (DOGE) has reduced the federal workforce by almost 280,000 employees, a central pledge of the current administration.
This action has generated significant apprehension among federal workers in regards to mental health and job security.
To understand the impact of DOGE's actions on federal worker perceptions of job security, we labeled 400 Reddit comments on topics related to the current reduction in federal workforce by DOGE as whether the author favored, opposed, or had a neutral stance.
We used these labels to build supervised learning models to predict stance.
Additionally, we employed unsupervised large language models (LLMs) to detect the stance of these Reddit comments from text, to further explore appropriate models for this current topic.

@@

We collected a total of 12,553 Reddit comments in early March 2024 from subreddits relevant to the topic of interest.
These comments were retrieved using keyword-based queries designed to capture posts of discussions that would reflect users’ stances on the federal workforce reduction by DOGE.
To prepare the data for analysis, a comprehensive preprocessing pipeline was conducted in Python 3.
This process involved removing HTML tags and web-URLs, stripping special characters and tagging artifacts such as “\@”, and replacing emojis with textual descriptions.
Also, it entailed eliminating punctuations and converting numerical values into their spelled-out forms.
The text was standardized to lowercase, lemmatized using the SpaCy library, and further refined by removing stopwords per the Natural Language Toolkit (NLTK) lists.
Spelling corrections were performed to improve textual coherence, while duplicate comments were identified and removed to maintain data integrity.

@@

All models were trained using default hyperparameter configurations.
The number of neighbors for KNN model was set to three, and the mtry parameter for Random Forrest was fixed at two, reflecting the relatively small number of metadata features in those models.
All experiments were conducted on a personal computer without the use of a GPU

In addition to traditional supervised models, we employed two instruction-tuned LLMs to perform zero- shot classification: Gemma 3.12B, developed by Google and based on the Gemini framework, and LLaMA 3.2B, developed by Meta AI. Gemma was deployed using the Great Lakes High-Performance Computing (HPC) cluster, while LLaMA was deployed locally on a GPU-enabled personal workstation using the rollama r-pacakge.
These models were selected for their strong zero-shot classification capabilities and ability to be run locally or on cluster environments without reliance on commercial APIs.
Both models were given the same prompt and tasks, which were developed by the research team to assess the effectiveness of classifying stance within our topic of interest (see Table 2).

@@

|  |  |
|--------------------|----------------------------------------------------|
| **Prompt**: | "Is this comment in 'favor', 'neutral', or 'oppose' the reduction in federal workforce? Provide one word answer only! |
| **Task**: | "You have assumed the role of a stakeholder that is presented with a Reddit comment from a likely federal worker related to the current policies on reducing the federal workforce. Please determine the author's stance on this topic, and provide the answer only." |

@@

```{r}
#| fig-width: 7
#| fig-height: 5
#| 
# number of total posts
ml_results |> 
  # filter models needed
  filter(model %in% 
           c("RandomForest", "KNeighbors"),
         metric %in% c("f1-score", "recall")) |> 
  ggplot(aes(x=model, y=score, fill=class)) +
  geom_col(position="dodge") +  
  coord_flip() +
  facet_grid(~metric) +
  guides(fill=guide_legend(title="", reverse = TRUE)) +
  scale_fill_viridis_d(option="F", direction = -1,
                       alpha=.85, end =.70, begin=.15) +
  labs(
    title = "Figure 1. Recall and Precision Scores for Random Forest
    and KNN Models",
    x = "Models",
    y = "Performance Scores"
  ) +
  theme_hc()

@@

Gemma and Llama demonstrated only fair agreement in stance classification (Cohen's Kappa = .24).
The contingency table ( see Table 2) reveals the highest agreement for "oppose" (.80), indicating some consistency in identifying a strong negative stance.
However, the agreement was substantially lower for "neutral" (.03) and "favor" (.01), highlighting the divergent interpretations of more nuanced language.
The off-diagonal values further illustrates these discrepancies, suggesting the fundamental differences in how the models process ambiguous cues and establish classification boundaries, particularly for less extreme stances.

@@

Public sentiment on Reddit were explored in terms of the DOGE federal workforce reduction through supervised learning (KNN, Random Forest) and unsupervised LLMs ( Gemma 3.12b and Llama 3.2 3B) for stance classifications.
Our supervised learning models overall achieved low to moderate success in this classification task, with the strongest performance in identifying opposing viewpoints, which were also the most prevalent in our labeled data.
Moreover, both models struggled with the "favor" and "neutral" stances, suggesting limitations in our features to capture these nuances.
Our unsupervised learning models also showed the best ability to classify "oppose" comments, yet exhibited significant challenges with the "favor" and "neutral" categories, particularly demonstrating low precision.

>>>>

---
title: "DOGE's Downsizing, Can AI Read the Reddit Room?"
format:
  jasa-pdf:
    keep-tex: true  
    journal:
      blinded: false
  jasa-html: default
date: last-modified
author:
  - name: Kevin Linares
    affiliations:
      - name: University of Maryland
  - name: Felix Baez-Santiago
  - name: Aria Lu
  - name: Gloria Zhou
    affiliations:
      - name: University of Michigan
abstract: |
  We investigated public sentiment on Reddit regarding the Department of Government Efficiency's federal workforce reduction by classifying 400 labeled comments (28% favor, 18% neutral, 54% oppose) using supervised and unsupervised Large Language Models (LLMs). It turned out that supervised models showed low-moderate success, particularly with "oppose" comments, but struggled with "favor" and "neutral" stances. Similarly, LLMs best intent identified "oppose" sentiment but exhibited low precision for "favor" and "neutral." These findings highlight the challenges of accurately gauging nuanced public opinion on government policy changes through automated methods on social media data.

@@

The newly formed Department of Government Efficiency (DOGE) has reduced the federal workforce by almost 280,000 employees, a central pledge of the current administration.
This action has generated significant apprehension among federal workers in regards to mental health and job security.
To understand the impact of DOGE's actions on federal worker perceptions of job security, we labeled 400 Reddit comments on topics related to the current reduction in workforce as whether the author favored, opposed, or had a neutral stance.
We used these labels to build supervised learning models to predict stance.
Additionally, we employed unsupervised large language models (LLMs) to detect the stance of these Reddit comments from text, to further explore appropriate models for this current topic.

@@

We collected 12,553 Reddit comments earlier this March from subreddits relevant to the topic of interest.
These comments were retrieved using keyword-based queries designed to capture discussion posts reflecting users’ stances on the federal workforce reduction by DOGE.
To prepare the data for analysis, a comprehensive preprocessing pipeline was conducted in Python 3.
This process involved removing HTML tags and web-URLs, stripping special characters and tagging artifacts such as “\@”, and replacing emojis with textual descriptions.
Also, it entailed eliminating punctuations and converting numerical values into their spelled-out forms.
The text was standardized to lowercase, lemmatized using the SpaCy library, and further refined by removing stopwords per the Natural Language Toolkit (NLTK) lists.
Spelling corrections were performed to improve textual coherence, while duplicate comments were identified and removed to maintain data integrity.

@@

All models were trained on a local machine with only the CPU using default hyperparameter configurations.
The number of neighbors for KNN model was set to three, and the mtry parameter for Random Forrest was fixed at two, reflecting the relatively small number of metadata features in those models.

We also employed two LLMs to perform zero-shot classification: Gemma 3.12B, developed by Google and based on the Gemini framework, and LLaMA 3.2B, developed by Meta AI. Gemma was deployed using the Great Lakes High-Performance Computing (HPC) cluster, while LLaMA was deployed locally on a dedicated-GPU using the rollama r-package.
These models were selected for their strong zero-shot classification capabilities and ability to be run locally or on cluster environments without reliance on commercial APIs.
Both models were given the same prompt and tasks, which were developed by the research team to assess the effectiveness of classifying stance within our topic (see Table 2).

@@

|  |  |
|----------------------|--------------------------------------------------|
| **Prompt**: | "Is this comment in 'favor', 'neutral', or 'oppose' the reduction in federal workforce? Provide one word answer only! |
| **Task**: | "You have assumed the role of a stakeholder that is presented with a Reddit comment from a likely federal worker related to the current policies on reducing the federal workforce. Please determine the author's stance on this topic, and provide the answer only." |

@@

```{r}
#| fig-width: 7
#| fig-height: 5.5
#| 
# number of total posts
ml_results |> 
  # filter models needed
  filter(model %in% 
           c("RandomForest", "KNeighbors"),
         metric %in% c("f1-score", "recall")) |> 
  ggplot(aes(x=model, y=score, fill=class)) +
  geom_col(position="dodge") +  
  coord_flip() +
  facet_grid(~metric) +
  guides(fill=guide_legend(title="", reverse = TRUE)) +
  scale_fill_viridis_d(option="F", direction = -1,
                       alpha=.85, end =.70, begin=.15) +
  labs(
    title = "Figure 1. Recall and Precision Scores for Random Forest
    and KNN Models",
    x = "Models",
    y = "Performance Scores"
  ) +
  theme_hc()

@@

Gemma and Llama demonstrated only fair agreement in stance classification (Cohen's Kappa = .24).
The contingency table ( see Table 3) reveals the highest agreement for "oppose" (.80), indicating some consistency in identifying a strong negative stance.
However, the agreement was substantially lower for "neutral" (.03) and "favor" (.01), highlighting the divergent interpretations of more nuanced language.
The off-diagonal values further illustrates these discrepancies, suggesting the fundamental differences in how the models process ambiguous cues and establish classification boundaries, particularly for less extreme stances.

@@

We explored public sentiment of DOGE's federal workforce reduction through supervised learning (KNN, Random Forest) and unsupervised LLMs ( Gemma 3.12b and Llama 3.2 3B) for stance classifications.
Our supervised learning models achieved low-moderate success in this classification task, with the strongest performance in identifying opposing viewpoints.
Moreover, both models struggled with the "favor" and "neutral" stances, suggesting limitations in our features to capture these nuances.
Our unsupervised learning models also showed the best ability to classify "oppose" comments, yet exhibited significant challenges with the "favor" and "neutral" categories, particularly demonstrating low precision.

### 2025-04-10 code/ml_llm/R/assign_3.tex
3rd draft of assing 3

\date{April 9, 2025}
\title{\bf DOGE's Downsizing, Can AI Read the Reddit Room?}
\author{
Kevin Linares\\
University of Maryland\\
and\\Felix Baez-Santiago\\
and\\Aria Lu\\
and\\Gloria Zhou\\
University of Michigan\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
We investigated public sentiment on Reddit regarding the Department of
Government Efficiency's federal workforce reduction by classifying 400
labeled comments (28\% favor, 18\% neutral, 54\% oppose) using
supervised and unsupervised Large Language Models (LLMs). It turned out
that supervised models showed low to moderate success, particularly with
``oppose'' comments, but struggled with ``favor'' and ``neutral''
stances. Similarly, LLMs best intent identified ``oppose'' sentiment but
exhibited low precision for ``favor'' and ``neutral.'' These findings
highlight the challenges of accurately gauging nuanced public opinion on
government policy changes through automated methods on social media
data.
\end{abstract}

@@

The newly formed Department of Government Efficiency (DOGE) has reduced
the federal workforce by almost 280,000 employees, a central pledge of
the current administration. This action has generated significant
apprehension among federal workers in regards to mental health and job
security. To understand the impact of DOGE's actions on federal worker
perceptions of job security, we labeled 400 Reddit comments on topics
related to the current reduction in federal workforce by DOGE as whether
the author favored, opposed, or had a neutral stance. We use these
labels to build supervised learning models to predict stance.
Additionally, we employ unsupervised large language models (LLMs) to
detect the stance of these Reddit comments from text, to further explore
appropriate models for this current topic.

@@

We collected a total of 12,553 Reddit comments in early March 2024 from
subreddits relevant to the topic of interest. These comments were
retrieved using keyword-based queries designed to capture posts of
discussions that would reflect users' stances on the federal workforce
reduction by DOGE. To prepare the data for analysis, a comprehensive
preprocessing pipeline was conducted in Python 3. This process involved
removing HTML tags and web-URLs, stripping special characters and
tagging artifacts such as ``@'', and replacing emojis with textual
descriptions. Also, it entailed eliminating punctuations and converting
numerical values into their spelled-out forms. The text was standardized
to lowercase, lemmatized using the SpaCy library, and further refined by
removing stopwords per the Natural Language Toolkit (NLTK) lists.
Spelling corrections were performed to improve textual coherence, while
duplicate comments were identified and removed to maintain data
integrity.

@@

All models were trained using default hyperparameter configurations. The
number of neighbors for KNN model was set to three, and the mtry
parameter for Random Forrest was fixed at two, reflecting the relatively
small number of metadata features in those models. All experiments were
conducted on a personal computer without the use of a GPU

In addition to traditional supervised models, we employed two
instruction-tuned LLMs to perform zero- shot classification: Gemma
3.12B, developed by Google and based on the Gemini framework, and LLaMA
3.2B, developed by Meta AI. Gemma was deployed using the Great Lakes
High-Performance Computing (HPC) cluster, while LLaMA was deployed
locally on a GPU-enabled personal workstation using the rollama
r-pacakge. These models were selected for their strong zero-shot
classification capabilities and ability to be run locally or on cluster
environments without reliance on commercial APIs. Both models were given
the same prompt and tasks, which were developed by the research team to
assess the effectiveness of classifying stance within our topic of
interest (see Table 2).

\hfill\break

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2778}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7222}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Prompt}: & ``Is this comment in `favor', `neutral', or `oppose'
the reduction in federal workforce? Provide one word answer only! \\
\textbf{Task}: & ``You have assumed the role of a stakeholder that is
presented with a Reddit comment from a likely federal worker related to
the current policies on reducing the federal workforce. Please determine
the author's stance on this topic, and provide the answer only.'' \\
\end{longtable}

@@

Gemma and Llama demonstrated only fair agreement in stance
classification (Cohen's Kappa = .24). The contingency table ( see Table
3) reveals the highest agreement for ``oppose'' (.80), indicating some
consistency in identifying a strong negative stance. However, agreement
was substantially lower for ``neutral'' (.03) and ``favor'' (.01),
highlighting the divergent interpretations of more nuanced language. The
off-diagonal values further illustrates these discrepancies, suggesting
the fundamental differences in how the models process ambiguous cues and
establish classification boundaries, particularly for less extreme
stances.

@@

Public sentiment on Reddit were explored in terms of the DOGE federal
workforce reduction through supervised learning (KNN, Random Forest) and
unsupervised LLMs ( Gemma 3.12b and Llama 3.2 3B) for stance
classifications. Our supervised learning models overall achieved low to
moderate success in this classification task, with the strongest
performance in identifying opposing viewpoints, which were also the most
prevalent in our labeled data. Moreover, both models struggled with the
``favor'' and ``neutral'' stances, suggesting limitations in our
features to capture these nuances. Our unsupervised learning models also
showed the best ability to classify ``oppose'' comments, yet exhibited
significant challenges with the ``favor'' and ``neutral'' categories,
particularly demonstrating low precision.

>>>>

\date{April 10, 2025}
\title{\bf DOGE's Downsizing, Can AI Read the Reddit Room?}
\author{
Kevin Linares\\
University of Maryland\\
and\\Felix Baez-Santiago\\
and\\Aria Lu\\
and\\Gloria Zhou\\
University of Michigan\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
We investigated public sentiment on Reddit regarding the Department of
Government Efficiency's federal workforce reduction by classifying 400
labeled comments (28\% favor, 18\% neutral, 54\% oppose) using
supervised and unsupervised Large Language Models (LLMs). It turned out
that supervised models showed low-moderate success, particularly with
``oppose'' comments, but struggled with ``favor'' and ``neutral''
stances. Similarly, LLMs best intent identified ``oppose'' sentiment but
exhibited low precision for ``favor'' and ``neutral.'' These findings
highlight the challenges of accurately gauging nuanced public opinion on
government policy changes through automated methods on social media
data.
\end{abstract}

@@

The newly formed Department of Government Efficiency (DOGE) has reduced
the federal workforce by almost 280,000 employees, a central pledge of
the current administration. This action has generated significant
apprehension among federal workers in regards to mental health and job
security. To understand the impact of DOGE's actions on federal worker
perceptions of job security, we labeled 400 Reddit comments on topics
related to the current reduction in workforce as whether the author
favored, opposed, or had a neutral stance. We used these labels to build
supervised learning models to predict stance. Additionally, we employed
unsupervised large language models (LLMs) to detect the stance of these
Reddit comments from text, to further explore appropriate models for
this current topic.

@@

We collected 12,553 Reddit comments earlier this March from subreddits
relevant to the topic of interest. These comments were retrieved using
keyword-based queries designed to capture discussion posts reflecting
users' stances on the federal workforce reduction by DOGE. To prepare
the data for analysis, a comprehensive preprocessing pipeline was
conducted in Python 3. This process involved removing HTML tags and
web-URLs, stripping special characters and tagging artifacts such as
``@'', and replacing emojis with textual descriptions. Also, it entailed
eliminating punctuations and converting numerical values into their
spelled-out forms. The text was standardized to lowercase, lemmatized
using the SpaCy library, and further refined by removing stopwords per
the Natural Language Toolkit (NLTK) lists. Spelling corrections were
performed to improve textual coherence, while duplicate comments were
identified and removed to maintain data integrity.

@@

All models were trained on a local machine with only the CPU using
default hyperparameter configurations. The number of neighbors for KNN
model was set to three, and the mtry parameter for Random Forrest was
fixed at two, reflecting the relatively small number of metadata
features in those models.

We also employed two LLMs to perform zero-shot classification: Gemma
3.12B, developed by Google and based on the Gemini framework, and LLaMA
3.2B, developed by Meta AI. Gemma was deployed using the Great Lakes
High-Performance Computing (HPC) cluster, while LLaMA was deployed
locally on a dedicated-GPU using the rollama r-package. These models
were selected for their strong zero-shot classification capabilities and
ability to be run locally or on cluster environments without reliance on
commercial APIs. Both models were given the same prompt and tasks, which
were developed by the research team to assess the effectiveness of
classifying stance within our topic (see Table 2).

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.3056}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.6944}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Prompt}: & ``Is this comment in `favor', `neutral', or `oppose'
the reduction in federal workforce? Provide one word answer only! \\
\textbf{Task}: & ``You have assumed the role of a stakeholder that is
presented with a Reddit comment from a likely federal worker related to
the current policies on reducing the federal workforce. Please determine
the author's stance on this topic, and provide the answer only.'' \\
\end{longtable}

@@

Gemma and Llama demonstrated only fair agreement in stance
classification (Cohen's Kappa = .24). The contingency table ( see Table
3) reveals the highest agreement for ``oppose'' (.80), indicating some
consistency in identifying a strong negative stance. However, the
agreement was substantially lower for ``neutral'' (.03) and ``favor''
(.01), highlighting the divergent interpretations of more nuanced
language. The off-diagonal values further illustrates these
discrepancies, suggesting the fundamental differences in how the models
process ambiguous cues and establish classification boundaries,
particularly for less extreme stances.

@@

We explored public sentiment of DOGE's federal workforce reduction
through supervised learning (KNN, Random Forest) and unsupervised LLMs (
Gemma 3.12b and Llama 3.2 3B) for stance classifications. Our supervised
learning models achieved low-moderate success in this classification
task, with the strongest performance in identifying opposing viewpoints.
Moreover, both models struggled with the ``favor'' and ``neutral''
stances, suggesting limitations in our features to capture these
nuances. Our unsupervised learning models also showed the best ability
to classify ``oppose'' comments, yet exhibited significant challenges
with the ``favor'' and ``neutral'' categories, particularly
demonstrating low precision.
